---
title: "R Notebook"
output: html_notebook
---

```{r setup}
library(tidyverse)
# library(signal)
library(plotly)
library(patchwork)
library(ggbeeswarm)
```

This is a script with the kinematics processing code.
```{r}
source('process_midlines.R')
```

Directory where all the tracked data files are stored.
```{r}
boxdir <- '/Users/etytel01/Box/Katz lamprey kinematics/'
datadir <- file.path(boxdir,'2021/Processed Data')
```

Names of the treatments.
```{r}
treatments = c("Control", "2hpi", "1dpi", "2dpi", "3dpi")
```

Body parts are named in DeepLabCut, but they have an order along the body, given here.
```{r}
bodypartorder = c("snout", "gill", "mid1", "mid2", "anus", "mid4", "tailbase", "tailtip")
```

Filming speed (frames per second)
```{r}
fps = 60
```

Load in approximate widthdata for ammocoetes, which we'll use for locating the center of mass.
```{r}
widthdata <- read_csv(file.path(boxdir,'2021/fishwidth.csv'))
```

Digitized scales for each day of filming.
```{r}
scalefile <- file.path(boxdir,'2021/Processed Data/scales.csv')
```

```{r}
scales <- read_csv(scalefile, col_types = list(Date = col_date("%m/%d/%y"),
                                               Animal = col_character(),
                                               Treatment = col_factor(levels = treatments)))
scales
```

# Load the DeepLabCut data and process it

```{r}
datafilenames <- list.files(datadir, pattern='*.csv', full.names = TRUE, recursive = TRUE)
datafilenames <- datafilenames[str_detect(datafilenames, 'shuffle1')]
```

```{r}
datafilenames <- datafilenames[str_detect(datafilenames, 'A1')]
```

## Load and process the data

Note that these chunks don't run by default, since they take a long time.

```{r eval = FALSE}
#options(error = recover)
data <- purrr::map_dfr(datafilenames, ~ load_katz_data(.x, widthdata = widthdata$ammowidth, fps = fps,
                 bodypartorder = bodypartorder, showfile = TRUE))
write_csv(data, file.path(boxdir,'Processed Data/A1data.csv')
```

This loads the previous processed data file, as produced above.
```{r}
data <- read_csv(file.path(boxdir,'2021/Processed Data/A1data.csv'))
```

## Look for bad frames

Look for variation in length
```{r}
animallengths <-
  data %>%
  get_arc_length() %>%
  group_by(Animal, trial, frame) %>%
  summarize(lenfr = last(s)) %>%
  ungroup() %>%
  group_by(Animal) %>%
  mutate(len = median(lenfr)) %>%
  ungroup()
```

```{r}
data <-
  data %>%
  left_join(animallengths, by = c("Animal","trial","frame"))
```

This also takes a long time
```{r eval=FALSE}
data <-
  data %>%
  get_point_order()
write_csv(data, file.path(boxdir,'2021/Processed Data/A1data.csv'))
```

```{r}
data <- read_csv(file.path(boxdir,'2021/Processed Data/A1data.csv'), 
                 col_types = list(Animal = col_character(),
                                  Treatment = col_factor(levels = treatments),
                                  bodyparts = col_factor(levels = bodypartorder)))
```

```{r}
data <-
  data %>%
  group_by(trial, frame) %>%
  mutate(pointorder = seq_len(n())) %>%
  mutate(goodpointorder = all(nextpoint == pointorder+1, na.rm=TRUE),
            goodlength = all(abs(lenfr - len) / len < 0.1))
```

## Process the kinematics for the good frames

```{r eval = FALSE}
data <-
  data %>%
  ungroup() %>%
  select(-(dx:wavelength)) %>%
   filter(goodpointorder & goodlength) %>%
  group_by(filename) %>%
  group_modify(~ get_all_kinematics(.x, widthdata = widthdata$ammowidth, fps = fps))
write_csv(data, file.path(boxdir,'2021/Processed Data/A1data-goodframes.csv'))
```

```{r}
data %>%
  ungroup() %>%
  select(-(dx:wavelength)) %>%
  filter(goodpointorder & goodlength) %>%
  filter(trial == '062921_A1V3_02') %>%
  get_arc_length() %>%
  interpolate_width(widthdata$ammowidth) %>%
  get_center_of_mass() %>%
  get_swim_vel_dir() %>%
  plot_ly(x = ~t, y = ~swimvelx, type = 'scatter', mode = 'markers')
    # get_swim_vel_dir()
```



# Plot kinematics

```{r}
data <- read_csv(file.path(boxdir,'2021/Processed Data/A1data-goodframes.csv'),
                 col_types = list(Date = col_date(), 
                                  Animal = col_character(),
                                  Treatment = col_factor(levels = treatments)))
```

```{r}
data %>%
  filter(trial == '062921_A1V3_02') %>%
  plot_ly(x = ~t, y = ~swimvel, color = ~trial, type = 'scatter', mode = 'lines') %>%
  layout(showlegend = FALSE)
```

