---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup}
library(tidyverse)
# library(signal)
library(plotly)
library(patchwork)
library(ggbeeswarm)
library(hdf5r)
```

This is a script with the kinematics processing code.
```{r}
source('process_midlines.R')
source('check_midlines.R')
```

Directory where all the tracked data files are stored.
```{r}
boxdir <- '/Users/etytel01/Box/Katz lamprey kinematics'
datadir <- file.path(boxdir,'2021/Processed Data/Exp_100921')
```

Names of the treatments.
```{r}
treatments = c("Control", "2hpi", "1dpi", "2dpi", "3dpi")
```

Body parts are named in DeepLabCut, but they have an order along the body, given here.
```{r}
bodypartorder = c("snout", "gill", "mid1", "mid2", "anus", "mid4", "tailbase", "tailtip")
```

Filming speed (frames per second)
```{r}
fps = 60
```

Load in approximate widthdata for ammocoetes, which we'll use for locating the center of mass.
```{r}
widthdata <- read_csv(file.path(boxdir,'2021/fishwidth.csv'))
```

Digitized scales for each day of filming.
```{r}
scalefile <- file.path(boxdir,'2021/Processed Data/scales.csv')
```

```{r}
scales <- read_csv(scalefile, col_types = list(Date = col_date("%Y-%m-%d"),
                                               Animal = col_character(),
                                               Treatment = col_factor(levels = treatments)))
scales
```

# Load the DeepLabCut data and process it

First run the ipython notebook "process_midlines.ipynb" to convert the DeepLabCut .h5 files into .csv files. (NB: I process the .h5 files a bit differently than DeepLabCut does to get the CSV files)
```{r}
datafilenames <- list.files(datadir, pattern='*.csv', full.names = TRUE, recursive = TRUE)
datafilenames <- datafilenames[str_detect(datafilenames, 'shuffle1')]
```

```{r}
datafilenames <- datafilenames[str_detect(datafilenames, '100921')]
```

## Load and process the data

Note that these chunks don't run by default, since they take a long time.


```{r eval = FALSE}
#options(error = recover)
data <- purrr::map_dfr(datafilenames, ~ load_katz_data(.x, scales,
                                                       widthdata = widthdata$ammowidth, fps = fps,
                 bodypartorder = bodypartorder, showfile = TRUE))
write_csv(data, file.path(boxdir,'2021/Processed Data/A21009data.csv'))
```

This loads the previous processed data file, as produced above.
```{r}
data <- read_csv(file.path(boxdir,'2021/Processed Data/A21009data.csv'),
                                  col_types = list(Animal = col_character(),
                                  Treatment = col_factor(levels = treatments),
                                  bodyparts = col_factor(levels = bodypartorder)))
```

## Look for bad frames

```{r}
data %>%
  ungroup() %>%
  distinct(trial)
```

First remove outliers based on a median filter procedure (after [https://ocefpaf.github.io/python4oceanographers/blog/2015/03/16/outlier_detection/])
```{r}
data.nooutliers <- data %>%
  group_by(Animal, trial) %>%
  group_modify(~ detect_and_replace_outliers(.x, k = 7, 
                                        outlierthreshold = 20,
                                        likelihoodthreshold = 0.4))
```


```{r}
data.nooutliers %>%
  filter(frame %% 5 == 0) %>%
  #filter(trial == '100921_A2V6_04' | trial == '100921_A2V6_13') %>%
  #filter(trial == '100921_A2V6_13') %>%
  group_by(trial, frame) %>%
  #filter(bodyparts == 'tailbase') %>%
  ggplot(aes(x = xmm, y = ymm)) +
  geom_line(aes(group = frame, color = Treatment)) +
  geom_point(data = ~filter(.x, isoutlier), aes(x = xmm0, y = ymm0), size=0.5, color='red') +
  facet_wrap(~trial, scales = 'free') +
  # scale_color_viridis_c() + #limits = c(0, 20)) +
  theme(
    # strip.background = element_blank(),
    # strip.text.x = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank()
  )
```

Look for variation in length
```{r}
animallengths <-
  data.nooutliers %>%
  get_arc_length() %>%
  group_by(Animal, trial, frame) %>%
  summarize(lenfr = last(s)) %>%
  ungroup() %>%
  group_by(Animal) %>%
  mutate(len = median(lenfr)) %>%
  ungroup()
```

```{r}
data.nooutliers <-
  data.nooutliers %>%
  select(-any_of(c('len', 'lenfr'))) %>%
  left_join(animallengths, by = c("Animal","trial","frame"))
```

This also takes a long time
```{r eval=FALSE}
data.nooutliers <-
  data.nooutliers %>%
  get_point_order()
write_csv(data.nooutliers, file.path(boxdir,'2021/Processed Data/A21009data.csv'))
```

```{r}
data <- read_csv(file.path(boxdir,'2021/Processed Data/A21009data.csv'), 
                 col_types = list(Animal = col_character(),
                                  Treatment = col_factor(levels = treatments),
                                  bodyparts = col_factor(levels = bodypartorder)))
```

```{r}
data <-
  data %>%
  group_by(trial, frame) %>%
  mutate(pointorder = seq_len(n())) %>%
  mutate(goodpointorder = all(nextpoint == pointorder+1, na.rm=TRUE),
            goodlength = all(abs(lenfr - len) / len < 0.1))
```

## Process the kinematics for the good frames

```{r eval = FALSE}
if ('dx' %in% colnames(data)) {
  data <- data %>%
    select(-dx:wavelength)
}
data <-
  data %>%
  ungroup() %>%
  filter(goodpointorder & goodlength) %>%
  group_by(filename) %>%
  group_modify(~ get_all_kinematics(.x, widthdata = widthdata$ammowidth))
write_csv(data, file.path(boxdir,'2021/Processed Data/A21009data-goodframes.h5'))
```


# Plot kinematics

```{r}
data <- read_csv(file.path(boxdir,'2021/Processed Data/A21009data-goodframes.h5'),
                 col_types = list(Animal = col_character(),
                                  Treatment = col_factor(levels = treatments),
                                  bodyparts = col_factor(levels = bodypartorder),
                                  curve = col_number()))
```

```{r}
data %>%
  ungroup() %>%
  distinct(trial)
```

```{r}
data1 <-
  data %>%
  filter(trial == '100921_A2V6_07')
```

```{r}
p1 <-
  data1 %>%
  filter((frame >= 20) & (frame <= 30)) %>%
  ggplot(aes(x = xmm - comx, y = ymm - comy, color = frame)) +
  geom_path(aes(group = frame)) +
  coord_equal()

p2 <- data1 %>%
  filter((frame >= 20) & (frame <= 30)) %>%
  ggplot(aes(x = s, y = curve, color = frame)) +
  geom_path()

p1 / p2
```

```{r}
data %>%
  ggplot(aes(x = t, y = swimvel, color = Treatment)) +
  geom_line(aes(group = trial))
```

```{r}
tailbeatdata <-
  data %>%
  arrange(trial, frame, bodyparts) %>%
  group_by(trial, frame) %>%
  summarize(across(c(freq,swimvel, Treatment, len, amp, wavespeed, wavelength, cycle_num), ~ last(na.omit(.x))),
            curve.mn = mean(curve, na.rm=TRUE)) %>%
  group_by(trial) %>%
  mutate(ncycles = max(cycle_num, na.rm = TRUE),
         iszerofreq = any(freq == 0, na.rm = TRUE)) %>%
  ungroup()
```


```{r}
tailbeatdata %>%
  filter(ncycles > 3 & !iszerofreq) %>%
  ggplot(aes(x = freq, y = swimvel, color = Treatment)) +
  geom_point() +
  facet_wrap(~ Treatment)
  # layout(showlegend = FALSE)
```
```{r}
tailbeatdata <-
  tailbeatdata %>%
  filter(ncycles > 3 & !iszerofreq)
```


```{r}
tailbeatdata %>%
  mutate(stridelen = swimvel/len / freq) %>%
  ggplot(aes(x = swimvel, y = stridelen, color = Treatment)) +
  geom_point() +
  facet_wrap(~Treatment)
```

```{r}
tailbeatdata %>%
  ggplot(aes(x = swimvel, y = amp, color = Treatment)) +
  geom_point() +
  facet_wrap(~Treatment)
```
```{r}
tailbeatdata %>%
  mutate(St = 2*freq*amp / swimvel) %>%
  ggplot(aes(x = swimvel, y = St, color = Treatment)) +
  geom_point() +
  facet_wrap(~Treatment) +
  ylim(0,1)
```
```{r}
tailbeatdata %>%
  ggplot(aes(x = swimvel, y = wavespeed, color = Treatment)) +
  geom_point() +
  facet_wrap(~Treatment) +
  ylim(0, 400)
```

```{r}
tailbeatdata %>%
  ggplot(aes(x = swimvel, y = wavelength, color = Treatment)) +
  geom_point() +
  facet_wrap(~Treatment)
```

```{r}
tailbeatdata2 <-
  tailbeatdata %>%
  filter(freq > 0.1 & freq < 5) %>%
  group_by(Treatment, trial, frame) %>%
  summarize(across(c(freq, amp, swimvel, wavespeed, wavelength, len), ~ median(.x, na.rm = TRUE))) %>%
  mutate(St = 2*freq*amp / swimvel,
         stridelen = swimvel/len / freq,
         slip = swimvel / wavespeed)
```

```{r}
p1 <- ggplot(tailbeatdata2, aes(x = Treatment, y = stridelen, color = Treatment)) +
  geom_violin() +
  geom_boxplot(width = 0.1) +
  stat_summary(aes(group = 1), geom = "line", fun = median) +
  ylim(0, 1)

p2 <- ggplot(tailbeatdata2, aes(x = Treatment, y = St, color = Treatment)) +
  geom_violin() +
  geom_boxplot(width = 0.1) +
  stat_summary(aes(group = 1), geom = "line", fun = median) +
  ylim(0, 1)

p3 <- ggplot(tailbeatdata2, aes(x = Treatment, y = slip, color = Treatment)) +
  geom_violin() +
  geom_boxplot(width = 0.1) +
  stat_summary(aes(group = 1), geom = "line", fun = median) +
  ylim(0,1)

p4 <- ggplot(tailbeatdata2, aes(x = Treatment, y = amp, color = Treatment)) +
  geom_violin() +
  geom_boxplot(width = 0.1) +
  stat_summary(aes(group = 1), geom = "line", fun = median)
  #ylim(0,1)

p5 <- ggplot(tailbeatdata2, aes(x = Treatment, y = swimvel, color = Treatment)) +
  geom_violin() +
  geom_boxplot(width = 0.1) +
  stat_summary(aes(group = 1), geom = "line", fun = median)
  #ylim(0,1)

p6 <- ggplot(tailbeatdata2, aes(x = Treatment, y = freq, color = Treatment)) +
  geom_violin() +
  geom_boxplot(width = 0.1) +
  stat_summary(aes(group = 1), geom = "line", fun = median)
  #ylim(0,1)

(p1 + p2 + p3 + p4 + p5 + p6) + plot_layout(guides = "collect")
```
```{r}
ampenvelope <-
  data %>%
  get_amplitude_envelope()
```

```{r}
ampenvelope
```

```{r}

ampenvelope %>%
  ggplot(aes(x = s.even, y = amp.even, color = Treatment, fill = Treatment)) +
  # geom_path(aes(group = paste(trial, cycle)), alpha = 0.2) +
  stat_summary(geom = 'ribbon', fun.data = 'mean_se', alpha = 0.5) +
  stat_summary(geom = 'path', fun = 'mean')
```