---
title: "R Notebook"
output: html_notebook
---

```{r setup}
library(tidyverse)
# library(signal)
library(plotly)
library(patchwork)
library(ggbeeswarm)
```

This is a script with the kinematics processing code.
```{r}
source('process_midlines.R')
```

Directory where all the tracked data files are stored.
```{r}
datadir <- '/Users/etytel01/Box/Katz lamprey kinematics/2021/Processed Data'
```

Names of the treatments.
```{r}
treatments = c("Control", "2hpi", "1dpi", "2dpi", "3dpi")
```

Body parts are named in DeepLabCut, but they have an order along the body, given here.
```{r}
bodypartorder = c("snout", "gill", "mid1", "mid2", "anus", "mid4", "tailbase", "tailtip")
```

Filming speed (frames per second)
```{r}
fps = 60
```

Load in approximate widthdata for ammocoetes, which we'll use for locating the center of mass.
```{r}
widthdata <- read_csv('/Users/etytel01/Box/Katz lamprey kinematics/2021/fishwidth.csv')
```

Digitized scales for each day of filming.
```{r}
scalefile <- '/Users/etytel01/Box/Katz lamprey kinematics/2021/Processed Data/scales.csv'
```

```{r}
scales <- read_csv(scalefile, col_types = list(Date = col_date("%m/%d/%y"),
                                               Animal = col_character(),
                                               Treatment = col_factor(levels = treatments)))
scales
```

# Test on one file
```{r}
testfile = '/Users/etytel01/Box/Katz lamprey kinematics/2021/Processed Data/Exp_062921/Exp_062921/Animal1/2dpi/063021_A1V08_04DLC_resnet50_LesionsOct13shuffle1_1030000.csv'
```

```{r}
process_filepath(testfile)
```

Load in the data.
```{r}
data1 <- read_csv(testfile)
```

And add in the scale data to get positions in mm and time in seconds.
```{r}
data1 <-
  bind_cols(data1, process_filepath(testfile)) %>%
  left_join(scales, by = c('Animal', 'Treatment', 'Date')) %>%
  mutate(t = frame / fps,
         xmm = x * Scale,
         ymm = y * Scale,
         bodyparts = factor(bodyparts, levels = bodypartorder),
         bodypartord = as.numeric(bodyparts))

```

Pull out the arc length and then estimate the median length.
```{r}
data1 <- get_arc_length(data1)
len1 <- get_median_length(data1)
```

Use that to interpolate the width.
```{r}
data1 <- interpolate_width(data1, widthdata = widthdata$ammowidth)
data1
```
Now estimate center of mass and plot it to make sure it makes sense.
```{r}
data1 <-
  get_center_of_mass(data1)

data1 %>%
  filter(frame >= 10 & frame <= 20) %>%
  group_by(frame) %>%
  plot_ly(x = ~xmm, y = ~ymm, type = "scatter", mode = "lines") %>%
  add_markers(x = ~comx, y = ~comy)
```
Estimate swimming velocity and direction.
```{r}
data1 <-
  get_swim_vel_dir(data1, fps = fps)
```

Then get the excursion of the body relative to the swimming direction.
```{r}
data1 <-
  data1 %>%
  get_excursions()
```


Plot the excursions of the snout and tail base and tip to check.
```{r}
data1 %>%
  filter(bodyparts %in% c("snout", "tailbase", "tailtip")) %>%
  group_by(bodyparts) %>%
  plot_ly(x = ~t, y = ~exc, color = ~bodyparts, type = "scatter", mode = "lines")
  
```

## Phase analysis

Get cycles between zero crossings of the excursions
```{r}
data1 <- 
  data1 %>% get_cycles()
```

```{r}
data1
```
The cycle plot should increase by 1 each time the tail crosses zero excursion.
```{r}
data1 %>%
  filter(bodyparts %in% c("tailtip", "anus", "snout")) %>%
  group_by(bodyparts) %>%
  plot_ly(x = ~phase, y = ~exc, color = ~bodyparts, type ="scatter", mode = "lines") %>%
  add_lines(y = ~cycle)
```
## Amplitude

This should show the mean amplitude as a function of position along the body.
```{r}
data1 %>%
  get_amplitudes() %>%
  group_by(bodypartord, cycle) %>%
  summarize(amp = first(na.omit(amp)),
            t = first(t)) %>%
  summarize(amp = mean(amp, na.rm = TRUE)) %>%
  ungroup() %>%
  plot_ly(x = ~bodypartord, y = ~amp, type = "scatter", mode = "lines+markers")
```

## Wave speed

And the wave speed as a function of position.
```{r}
data1 %>%
  get_wavespeed() %>%
  group_by(bodyparts, bodypartord) %>%
  summarize(wavespeed = mean(wavespeed, na.rm = TRUE)) %>%
  ungroup() %>%
  plot_ly(x = ~bodypartord, y = ~wavespeed, type = "scatter", mode = "lines+markers")
```
Usually I assume that wavespeed is constant along the body, but here it appears to be decreasing toward the tail. Could be that the wavelength is growing along the body.

## Wavelength

Estimate the wavelength.
```{r}
data1 %>%
  get_wavelength() %>%
  group_by(bodyparts, bodypartord) %>%
  summarize(wavelength = mean(wavelength, na.rm = TRUE)) %>%
  ungroup() %>%
  plot_ly(x = ~bodyparts, y = ~wavelength, type = "scatter", mode = "lines+markers")
  
```
It's pretty constant along the body where we can measure it, which is mostly the anterior body.


# Test on a second file

```{r}
testfile2 = '/Users/etytel01/Box/Katz lamprey kinematics/2021/Processed Data/Exp_062921/Exp_062921/Animal1/2hpi/062921_A1V14_04DLC_resnet50_LesionsOct13shuffle1_1030000.csv'
```

```{r}
data2 <-
  load_katz_data(testfile2, 
                 widthdata = widthdata$ammowidth, fps = fps,
                 bodypartorder = bodypartorder) %>%
  get_all_kinematics(widthdata = widthdata$ammowidth, fps = fps)
```
Summary plot
```{r}
p1 <- ggplot(data2, aes(x = t, y = swimvel)) +
  geom_line()

p2 <- 
  data2[seq(1,nrow(data2),30),] %>%
  ggplot(aes(x = t, y = 1, angle = atan2(swimdiry, swimdirx), radius = 0.05)) +
  geom_spoke() +
  coord_fixed()

p3 <-
  data2 %>%
  group_by(frame,t) %>%
  summarize(wavespeed = mean(wavespeed, na.rm = TRUE)) %>%
  ggplot(aes(x = t, y = wavespeed)) +
  geom_point()

p4 <-
  data2 %>%
  group_by(frame,t) %>%
  summarize(wavelength = last(na.omit(wavelength))) %>%
  ggplot(aes(x = t, y = wavelength)) +
  geom_point()

p5 <-
  data2 %>%
  group_by(bodyparts, bodypartord) %>%
  summarize(amp = mean(amp, na.rm = TRUE),
            s = mean(s, na.rm = TRUE)) %>%
  ggplot(aes(x = s, y = amp)) +
  geom_line()
  #geom_point(aes(color = bodyparts))

p6 <-
  data2 %>%
  group_by(bodyparts, bodypartord) %>%
  summarize(wavelength = mean(wavelength, na.rm = TRUE),
            s = mean(s, na.rm = TRUE)) %>%
  ggplot(aes(x = s, y = wavelength)) +
  geom_line()
  #geom_point(aes(color = bodyparts))

p1 + p3 + p4 + p5 + p6
```

# Load many files

```{r}
datafilenames <- list.files(datadir, pattern='*.csv', full.names = TRUE, recursive = TRUE)
datafilenames <- datafilenames[str_detect(datafilenames, 'shuffle1')]
```

```{r}
datafilenames <- datafilenames[str_detect(datafilenames, 'A1')]
```

```{r eval = FALSE}
#options(error = recover)
A1data <- purrr::map_dfr(datafilenames[1:5], ~ load_katz_data(.x, widthdata = widthdata$ammowidth, fps = fps,
                 bodypartorder = bodypartorder, showfile = TRUE))
#write_csv(A1data, '/Users/etytel01/Box/Katz lamprey kinematics/2021/Processed Data/A1data.csv')
```

```{r}
A1data <-
  A1data %>%
  group_by(filename) %>%
  group_modify(~ get_all_kinematics(.x, widthdata = widthdata$ammowidth, fps = fps))
```

```{r}
A1data <- read_csv('/Users/etytel01/Box/Katz lamprey kinematics/2021/Processed Data/A1data.csv')
```

Pull out the median length of each animal across all of the trials.
```{r}
animallengths <-
  A1data %>%
  group_by(Animal, trial, frame) %>%
  summarize(len = last(s)) %>%
  ungroup() %>%
  group_by(Animal) %>%
  summarize(len = median(len, na.rm = TRUE))
```

And merge it back into the main data set.
```{r}
A1data <-
  left_join(A1data, animallengths, by = "Animal")
```

```{r}
A1data <-
  A1data %>%
  group_by(trial, frame, t) %>%
  mutate(lenfr = last(s)) %>%
  ungroup()
```

```{r}
A1data %>%
    group_by(trial, frame, t) %>%
    summarize(lenfr = last(lenfr)) %>%
  plot_ly(x = ~trial, y = ~lenfr, color = ~trial, type="box",
          showlegend = FALSE)
```

```{r}

```

```{r}
testtrial <-
  A1data %>%
  filter(trial == '062921_A1V16_02')
```

```{r}
testtrialbyframe <-
  testtrial %>%
  group_by(frame,t) %>%
  summarize(across(c(comx,comy, swimvel, s), ~last(.x)))

p1 <- testtrialbyframe %>%
  ggplot(aes(x = t, y = swimvel)) +
  geom_line()
p2 <- testtrialbyframe %>%
  ggplot(aes(x = t, y = s)) +
  geom_line()

p1 + p2
```

```{r}
testtrial %>%
  group_by(frame) %>%
  plot_ly(x = ~xmm, y = ~ymm, frame = ~frame, type = "scatter", mode = "lines") %>%
  add_markers(color = ~-log10(likelihood))
```


```{r}
badtrials <-
  A1data %>%
  group_by(trial, frame) %>%
  summarize(lenfr = last(lenfr),
            len = last(len),
            lenerror = abs(lenfr - len)/len) %>%
  summarize(lenerror_mn = mean(lenerror, na.rm = TRUE),
    nbad = sum(lenerror > 0.1),
            badtrial = nbad > 10)

badtrials
```

```{r}
goodtrials <-
  badtrials %>%
  filter(!badtrial) %>%
  pull(trial)

goodtrials
```

```{r}
A1data <-
  A1data %>%
  filter(trial %in% goodtrials)
```

```{r}
p1 <- ggplot(A1data, aes(x = t, y = swimvel, color = trial)) +
  geom_line() +
  theme(legend.position = 'none')

p1
```

```{r}
summary_kinematics_plot <- function(df) {
  p1 <- ggplot(df, aes(x = t, y = swimvel, color = trial)) +
    geom_line()
  
  p2 <-
    df %>%
    group_by(trial, frame,t) %>%
    summarize(wavespeed = mean(wavespeed, na.rm = TRUE)) %>%
    ggplot(aes(x = t, y = wavespeed, color = trial)) +
    geom_point()
  
  p3 <-
    df %>%
    group_by(trial, frame,t) %>%
    summarize(wavespeed = mean(wavespeed, na.rm = TRUE),
              swimvel = first(swimvel)) %>%
    ggplot(aes(x = swimvel, y = wavespeed, color = trial)) +
    geom_point()

  p6 <-
    df %>%
    group_by(trial, frame,t) %>%
    summarize(wavelength = last(na.omit(wavelength))) %>%
    ggplot(aes(x = t, y = wavelength, color = trial)) +
    geom_point()
  
  p4 <-
    df %>%
    group_by(trial, bodyparts, bodypartord) %>%
    summarize(amp = mean(amp, na.rm = TRUE),
              s = mean(s, na.rm = TRUE)) %>%
    ggplot(aes(x = s, y = amp, color = trial)) +
    geom_line()
  #geom_point(aes(color = bodyparts))
  
  p5 <-
    df %>%
    group_by(trial, bodyparts, bodypartord) %>%
    summarize(wavelength = mean(wavelength, na.rm = TRUE),
              s = mean(s, na.rm = TRUE)) %>%
    ggplot(aes(x = s, y = wavelength, color = trial)) +
    geom_line()
  #geom_point(aes(color = bodyparts))
  
  p1 + p2 + p3 + p4 + p5 + p6 +  plot_layout(guides = 'none')
}
```

```{r}
summary_kinematics_plot(A1data)
```

```{r}
A1data %>%
  group_by(Treatment) %>%
  ggplot(aes(x = Treatment, y = swimvel)) +
  geom_violin() +
  stat_summary(fun.data = "mean_sdl")
```

